---
title: "Fall Chinook Tagging Summary 2016"
author: "Gabriel Singer"
date: "December 20, 2016"
output: word_document
---

```{r setup, include=FALSE}
library(readxl)
library(readr)
library(dplyr)
library(chron)
library(wesanderson)
library(stargazer)
library(knitr)
library(ggplot2)
library(devtools)

##Install R packages written by Myfanwy: fishtrackr, 
## run at beginning before sourcing csvs written ## 
# get.CDEC("LIS","E",25,"2015-09-05","2015-11-15") # get LIS sensor info for temp
#  cdec.dat$date <- as.POSIXct(cdec.dat$date, format = "%Y-%M-%D") # convert to usable format
# 
#  #clean up and write to csv
# dd <- cdec.dat %>% 
#   group_by(date) %>% 
#   summarise(meantemp = mean(sensor_25))
# which(dd$meantemp > 75) # figure out which values are weird
# dd[53, 2] <- NA # reassign weird values to NA
# dd$meantempC <- (dd$meantemp - 32)*(5/9) # convert to celcius
# write.csv(dd, "cdecdat.csv")

## Doing same for DO:
# get.CDEC("LIS","E", 61 ,"2015-09-05","2015-11-15")

# Load data
d <- read_excel("../data/AdultChinook_YBP_2015.xlsx")
dd <- read_csv("../data/cdecdat.csv")
dd$date <- as.POSIXct(dd$date, format = "%Y-%M-%D") # convert to usable format

f <- read_csv("../data/cdecdatDO.csv")
f$date <- as.POSIXct(f$date, format = "%Y-%M-%D")
f <- f %>% 
   group_by(date) %>% 
   summarise(meanDO = mean(sensor_61))

d <- select(d, -c(co2, floy, genetics, rampnotes))
d <- d[1:30,]
# Extract times only
d$toc <- as.POSIXlt(d$toc)
d$toc <- strftime(d$toc, format = "%H:%M:%S") 
d$toc <- times(d$toc) # This is what I was looking for

# Do the same for the other time columns:
d$surgstart <- as.POSIXlt(d$surgstart)
d$surgstart <- strftime(d$surgstart, format = "%H:%M:%S") 
d$surgstart <- times(d$surgstart) 
head(d$surgstart)

d$surgend <- as.POSIXlt(d$surgend)
d$surgend <- strftime(d$surgend, format = "%H:%M:%S") 
d$surgend <- times(d$surgend) 

d$tor <- as.POSIXlt(d$tor)
d$tor <- strftime(d$tor, format = "%H:%M:%S") 
d$tor <- times(d$tor)

# Add handling time column
d$handlingtime <- times(d$tor - d$toc, out.format = "h:m")
head(d$handlingtime)
class(d$handlingtime)
d$handlingtime <- as.numeric(as.character(d$handlingtime))

# Add surgery time column
d$surgtime <- times(d$surgend - d$surgstart, out.format = "h:m")
d$surgtime <- as.numeric(as.character(d$surgtime))
head(d$surgtime)

detach("package:chron", unload = TRUE)
```

## Basic Summaries

```{r kable_stargaze, results='asis'}
# Tagging locations
d5 <- d %>% 
  group_by(tagloc) %>% 
  summarise(N = n(), meanhandle = mean(handlingtime), meanrecov = mean(recov), meanramp = mean(c(TailGrab, BodyFlex, HeadComplex, VOR, Orientation)))
kable(d5, caption = "Breakdown By Tag Site", col.names = c("Tagging Location", "N", "Mean Handling Time", "Mean Recovery Time", "Mean Ramp Score"), digits = 2)

d2 <- 
  d %>% count(sex)
kable(d2, caption = 'Number of Males and Females', col.names = c("Sex", "N"))

d3 <- d %>% 
  group_by(surgeon) %>% 
  summarise(n = n(), mean_surgtime = mean(surgtime)) %>% 
  arrange(-n)
kable(d3, caption = "Surgery Summaries", col.names = c("Surgeon", "N", "Mean Surgery Time"), digits = 2)

```

### Demographic Summaries
```{r demplots, fig.height=5, fig.width=10}

## Set palette from here on out to Darjeeling:
dar <- wes_palette("Darjeeling", type = "discrete")
darc <- wes_palette("Darjeeling", type = "continuous")

# mean difference in fork length/total length between males and females?

ggplot(d, aes(x = sex, y = fl)) + 
  geom_boxplot(aes(color = sex, fill = sex, alpha = 0.8)) + 
  ylab("Fork Length (mm)") + scale_x_discrete(labels = c("Female", "Male", "Unknown")) + 
  ggtitle("Fork Length By Sex") + 
  theme(legend.position = "none") + xlab(" ") + scale_color_manual(values = wes_palette("Darjeeling")) + scale_fill_manual(values = wes_palette("Darjeeling"))

ggplot(d, aes(x = sex, y = tl)) + 
  geom_boxplot(aes(color = sex, fill = sex, alpha = 0.8)) + 
  ylab("Total Length (mm)") + scale_x_discrete(labels = c("Female", "Male", "Unknown")) + 
  ggtitle("Total Length By Sex") + 
  theme(legend.position = "none") + xlab(" ") + scale_color_manual(values = wes_palette("Darjeeling")) + scale_fill_manual(values = wes_palette("Darjeeling"))

```

## Seasonal Summaries

```{r plots, fig.height=5, fig.width=10, warning=FALSE, message=FALSE}

## Set palette from here on out to Darjeeling:
dar <- wes_palette("Darjeeling", type = "discrete")
darc <- wes_palette("Darjeeling", type = "continuous")

ggplot(d, aes(x = datetagged, y = handlingtime)) + geom_point(aes(color = tagloc), position = "jitter", size = 3) +
  ggtitle("Handling Times Across Tagging Season") + ylab("Handling Time (mins)") + xlab("Date") + geom_smooth(color = "black", size = 0.3) + scale_color_manual(values = dar)

# where did we catch fish over time?
d %>% 
  group_by(datetagged, tagloc) %>% 
  summarise(count = n()) %>% 
ggplot(., aes(x = datetagged, y = count)) + geom_bar(aes(fill = tagloc), stat = "identity", position = "dodge" ) + scale_fill_manual(values = dar) + ggtitle("Tagging Location and Numbers Across Time") + ylab("Number of Fish Tagged") + xlab("Date") + theme(legend.title = element_blank())

# what tide did we catch fish on?
d %>% 
  group_by(tide) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x = tide, y = count)) + geom_bar(aes(fill = tide, width = 0.5), stat = "identity") + theme(legend.position = "none") + ggtitle("More Fish Were Caught During Flood Tide \n (reflects documented catches, not total effort)") + scale_fill_manual(values = dar) + scale_x_discrete(labels = c("Ebb", "Flood", "Slack")) + xlab(" ") + ylab("Number of Fish")

# Toe Drain Temperature vs. Tagged Fish

ggplot(d, aes(x = datetagged, y = TD_temp)) + geom_point(size = 2) + geom_line(data = dd, aes(x = date, y = meantempC), color = "#00A08A") + ggtitle("Toe Drain Temperature Across Tagging Season") + ylab("Temperature (C)") + xlab("Date")

# Toe Drain DO vs. Tagged Fish
ggplot(d, aes(x = datetagged, y = TD_do)) + geom_point(size = 2) + geom_line(data = f, aes(x = date, y = meanDO), color = "#046C9A") + ggtitle("Toe Drain Dissolved Oxygen Across Tagging Season") + ylab("Dissolved Oxygen (mg/L)") + xlab("Date")
```
Black dots denote days when fish were tagged.  Note that the temperature and dissolved oxygen lines represent daily averages from the Lisbon Weir CDEC station; these were recorded recorded at the time of tagging as well, and are mapped by the black dots.

```{r DO_and_temp, fig.height=5, fig.width=10}
# Dissolved Oxygen and Temp vs. Ramp Score
d <- d %>% 
  group_by(tagid) %>% 
  mutate(meanramp = mean(c(TailGrab, BodyFlex, HeadComplex, VOR, Orientation)))

# Count number of fish tagged on each day and add as a new variable:
d <- d %>% 
  group_by(datetagged) %>% 
  mutate(count = n()) %>% 
  ungroup()

ggplot(d, aes(x = TD_temp, y = meanramp)) + geom_point(aes(size = count), color = "#046C9C", alpha = 0.6) + ggtitle("Mean Aggregate RAMP Score Increases with Temperature") + ylab("Aggregate RAMP Score") + xlab("Temperature (C)") + scale_color_manual(values = dar) + theme(legend.title = element_blank())

```

Blue dots denote days when fish were tagged; size of the dot indicates how many were tagged on that day with the same aggregate ramp score (possible values are 1:7).  As we can see, the highest RAMP scores (indicating the most impaired fish) were coincident with higher temperatures.  As such, they were also coincident with earlier dates of tagging (not plotted).

Banana results will need further analysis; but in general, we caught more fish on days when no bananas were eaten. :)




